---
title: "Figure 3 DGE Scatter & GSEA Bar plots for BMT Treg in Colon vs. non-Colon"
subtitle: "Figure 3 DGE Scatter comparing transplanted colon- vs. non-colon-homing Treg. GSEA Barplots for Genes up/down in colon-resident Treg for BMT Treg ranked according to colon vs. non-colon DGE comparison. Also, Supplementary Figure 3 subfigures based on the same subset of samples are generated."
author: "David Dittmar"
date: "`r Sys.Date()`"
bibliography: Rscripts.bib
nocite: '@MIRAGAIA2019493'
output:
  html_document:
    df_print: paged
    fig_caption: yes
    number_sections: no
    toc: yes
    toc_float: yes
    code_folding: hide 
---

***

```{r, include=FALSE}
# Load required libraries
library(edgeR)
library(AnnotationDbi)
library(org.Mm.eg.db) #for mouse
library(tibble)
library(ggplot2)
library(ggrepel)
```

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
# ...packages.bib file is also written here
knitr::write_bib(.packages(), "Figure3_TregDGEScatter_GSEA_cvnc.pc_BMT_packages.bib")
```

```{r}
# Load required counts, metadata and annotaion RDS files
counts.m.rm<-readRDS("./RCTDIR/counts.m.rm.2021-06-09.rds")
metad.m.rm<-readRDS("./METADIR/metad.m.rm.2021-06-09.rds")
stid<-readRDS("./METADIR/stid_2020-12-21.rds")
```
* First, required data objects (counts, metadata, short transcript ID) and R packages are loaded. 
* This is the starting point for most analyses conducted on bulk RNA-seq data for the associated manuscript.

***

# Data Preparation
* For the analysis, the data and metadata have to be prepared and properly annotated.

## Annotation Data Frame
* An annotation data frame with 4 columns is constructed.
* The additional EntrezID column is required for possible downstream pathway analyses or annotation operations.
```{r}
genes.df<-as.data.frame(strsplit2(stid,"[$]"))
colnames(genes.df)<-c("EnsemblID","GeneSymbol","Length","GeneType")
genes.df$EnsemblID<-as.character(genes.df$EnsemblID)
genes.df$GeneSymbol<-as.character(genes.df$GeneSymbol)
genes.df$Length<-as.numeric(as.character(genes.df$Length))
genes.df$GeneType<-as.character(genes.df$GeneType)
genes.df$EntrezID<- mapIds(org.Mm.eg.db,
                    keys=genes.df$GeneSymbol,
                    column="ENTREZID",
                    keytype="SYMBOL",
                    multiVals="first")
rownames(genes.df)<-stid
head(genes.df, n=5L)
#See if the gene length values are still as they should, so this should yield an empty vector
which(strsplit2(stid,"[$]")[,3] %in% genes.df$Length==FALSE)
```

## Subsetting Index
* In this analysis, we focus on donor Treg re-isolated from recipient organs 7d after transfer (BMT w/o Tconv & BMT w Tconv).
* A subsetting index is generated accordingly.
```{r}
cvnc.pc.ind<-(grepl("prc",metad.m.rm$appl)|grepl("pro",metad.m.rm$appl))&(grepl("S", metad.m.rm$org)|grepl("C", metad.m.rm$org)|grepl("L", metad.m.rm$org))
print(paste(sum(cvnc.pc.ind),"samples enter the analysis!"))
```

## Counts
* Here, the comprehensive counts data frame is reduced using the subsetting index.
```{r}
#Subsetting the respective counts object (counts.m.rm)
counts.cvnc.pc<-counts.m.rm[,cvnc.pc.ind]
paste(ncol(counts.cvnc.pc),"samples are selected:")
colnames(counts.cvnc.pc)
```

## Metadata
* The metadata has to be reduced accordingly. 
* The original order of samples is introduced for downstream re-ordering of the read count table.
* An additional group variable assigning colon-re-isolated samples to one group and spleen- or liver-migrated Treg to the other is introduced, as well.
* This enables the comparison of lymphoid vs. non-lymphoid tissue-homing Treg in the deployed BMT models, underlying this analysis.
```{r}
metad.cvnc.pc<-metad.m.rm[cvnc.pc.ind,]
metad.cvnc.pc$orig.ord<-1:nrow(metad.cvnc.pc)
metad.cvnc.pc<-metad.cvnc.pc[order(metad.cvnc.pc$grp),]
metad.cvnc.pc<-metad.cvnc.pc[order(metad.cvnc.pc$org),]
metad.cvnc.pc$org.alt<-c(rep("colon",nrow(metad.cvnc.pc[metad.cvnc.pc$org=="C",])),rep("spleen/liver",nrow(metad.cvnc.pc[metad.cvnc.pc$org!="C",])))
counts.cvnc.pc.ind<-metad.cvnc.pc$orig.ord
head(metad.cvnc.pc, n = 4L)
#the group variable for all later assignments of colors/pchs.
group.cvnc.pc<-factor(as.character(metad.cvnc.pc$org.alt))
group.cvnc.pc
```

## Re-ordering Counts Data Frame Columns
* Here, read counts are re-ordered according to the metadata object.
```{r}
counts.cvnc.pc.ord<-counts.cvnc.pc[,counts.cvnc.pc.ind]
head(counts.cvnc.pc.ord, n=4L)
counts.cvnc.pc<-counts.cvnc.pc.ord
rm(counts.cvnc.pc.ord)
```

## Consistency Check
* In order to avoid wrongful labelling, sample_IDs are used to confirm that the counts and metadata objects are consistent.
```{r}
if (sum(colnames(counts.cvnc.pc)==metad.cvnc.pc$sample_ID)==sum(cvnc.pc.ind)) {
  print("Counts and metadata are properly subsetted and in accordance with each other. The analysis can be proceeded with")
}else {
  print("Something is inconsistent, check the code above for errors during subsetting!")
}
```

## DGE List Object
* A DGEList object is generated, the main class of the edgeR package. 
* For full functionality, the counts.cvnc.pct, group.avp.cvnc.pct and genes.df variables are used to do so.
```{r}
dgel.cvnc.pc<-DGEList(counts = counts.cvnc.pc, group = group.cvnc.pc, genes = genes.df)
head(dgel.cvnc.pc$samples, n=4L)
head(dgel.cvnc.pc$genes, n=4L)
```

## Exclusion of Rare Transcripts
* The "filterByExpr()" function is used to determine which genes have sufficiently large counts to be retained in a statistical analysis. 
* Only genes fulfilling that criterion are kept and also normalization factors are calculated.
```{r}
keep.cvnc.pc <- filterByExpr(dgel.cvnc.pc)
dgel.cvnc.pc <- dgel.cvnc.pc[keep.cvnc.pc, , keep.lib.sizes=FALSE]
dgel.cvnc.pc <- calcNormFactors(dgel.cvnc.pc)
print(paste(table(keep.cvnc.pc)[1],"genes are discarded while",table(keep.cvnc.pc)[2], "genes are retained in the analysis!"))
```

***

# Differential Gene Expression (DGE) analysis
* Differential Gene Expression (DGE) is performed in the following part.

## Design for DGE
* Hereafter, a design is created that contains parameters that are of analytic interest as well as a blocking factor for the number of pcr2 cycles applied in library preparation.
```{r}
# Blocking Factor
pcr2<-factor(metad.cvnc.pc$PCR2.Cy)
# Design
cvnc.pc.design.dge<-model.matrix(~0+group.cvnc.pc+pcr2)
colnames(cvnc.pc.design.dge)<-c("colon", "spleen_liver", "twelve", "thirteen", "fourteen")
as_tibble(cvnc.pc.design.dge)
```

## Dispersion Esitmation
* Hereafter, dispersion is estimated.
```{r}
dgel.cvnc.pc<-estimateDisp(dgel.cvnc.pc,design=cvnc.pc.design.dge,robust=TRUE)
dgel.cvnc.pc$common.dispersion
```

## GLM Fitting
* Fitting gene wise glm (generalized linear model) for the constructed design.
```{r}
cvnc.pc.fit<-glmQLFit(dgel.cvnc.pc,cvnc.pc.design.dge)
```

## DGE for Contrast of Interest (COI)
* The interest in this analysis lies in the differences between Treg re-isolated from colon versus spleen&liver. 
* Thus, our contrast of interest is "colon - spleen_liver" for the DGE analysis. 
* We can have a quick look at how many genes are DE with a FC > 1.5 and FDR < 0.05.
```{r}
# Contrast of Interest
cvnc.pc.con<-makeContrasts(colon - spleen_liver, levels=cvnc.pc.design.dge) 
# DGE testing against fc = 1.5.
cvnc.pc.tr <- glmTreat(cvnc.pc.fit, contrast=cvnc.pc.con, lfc=log2(1.5))
cvnc.pc.tr.qstat<-topTags(cvnc.pc.tr,n = Inf)
cvnc.pc.tr.qdt<-decideTestsDGE(cvnc.pc.tr)
summary(cvnc.pc.tr.qdt)
```

### Exporting DGE Statistics
* The data frame contained in the glmTreat function results is stored for direct access.
```{r}
write.table(x = cvnc.pc.tr.qstat$table,
            file = "./ANALDIR/Figure3b_TregDGE_cvnc_pro.prc_FC1.5.txt",
            quote=FALSE, 
            sep = "\t",
            col.names = TRUE,
            row.names = FALSE)
```

***

# DGE Visualization
* Hereafter, DEGs are visualized *via* a scatter plot.
* In Addition, Genes up/down in colon-resident Cd62L- Treg will be ranked according to log2(FC) of the respective comparison in BMT Treg (w & w/o Tconv).

## Data Preparation
* The data is filtered based on logCPM and logRPKM thresholds.
* Necessary data transformations for the scatter plot are carried out as well: columns with individual logCPMs (calculated from logCPM and logFC)
```{r}
# Extraction of data frame from glmTreat result
cvnc.pc.tr.qstat.scat<-cvnc.pc.tr.qstat$table #is a data frame
# Introduction of logRPKM column
cvnc.pc.tr.qstat.scat$logRPKM <- cvnc.pc.tr.qstat.scat$logCPM - log2(cvnc.pc.tr.qstat.scat$Length * 0.001)
# Commonly used filter criteria
cvnc.pc.tr.qstat.scat.filt<-subset(cvnc.pc.tr.qstat.scat, logCPM > 1 & logRPKM > 1)
nrow(cvnc.pc.tr.qstat.scat.filt)
# Transformation for scatter plot
cvnc.pc.tr.qstat.scat.filt$logCPMnonColon<-cvnc.pc.tr.qstat.scat.filt$logCPM - (cvnc.pc.tr.qstat.scat.filt$logFC/2) #non-colon
cvnc.pc.tr.qstat.scat.filt$logCPMcolon<-cvnc.pc.tr.qstat.scat.filt$logCPM + (cvnc.pc.tr.qstat.scat.filt$logFC/2) #colon
# Separation of up and down genes
cvnc.pc.tr.qstat.scat.filt.UP<-subset(cvnc.pc.tr.qstat.scat.filt, logFC >= 0 & FDR < 0.05)
cvnc.pc.tr.qstat.scat.filt.DN<-subset(cvnc.pc.tr.qstat.scat.filt, logFC <= 0 & FDR < 0.05)
nrow(cvnc.pc.tr.qstat.scat.filt.UP)
nrow(cvnc.pc.tr.qstat.scat.filt.DN)
```

## Exporting UP and DN regulated Genes
* Significantly up and down regulated genes are exported in the following.
```{r}
#extract filtered gene lists
write.table(rownames(cvnc.pc.tr.qstat.scat.filt.UP),
            file = "./ANALDIR/Figure3b_TregDGE_cvnc_pro.prc.filt.sig.UP.txt",
            col.names = "GeneID",
            quote=FALSE, 
            row.names = FALSE)
write.table(rownames(cvnc.pc.tr.qstat.scat.filt.DN),
            file = "./ANALDIR/Figure3b_TregDGE_cvnc_pro.prc.filt.sig.DN.txt",
            col.names = "GeneID",
            quote=FALSE, 
            row.names = FALSE)
```

## Scatterplot

### Definition of Genes of Interest (GOI)
* Hereafter, GOI are taken from Miragaia et al. 2019 (Figure 2 B) presenting Treg subpopulations of the steady-state.
* The question addressed here is whether our data is in line with their findings. 
* [Here you can have a look at the figure](https://pubmed.ncbi.nlm.nih.gov/30737144/#&gid=article-figures&pid=figure-2-uid-2)
```{r}
# non-lymphoid tissue
tm_fig2_panel_NLT<-rev(c("Ccr1","Ccr2","Gzmb","Il10","Il1rl1","Kdm6b","Gata3","Areg","S100a4","Itgae","Rora","Klrg1","Lgals1","Fgl2","Cmtm7","Cd44","Icos","Ctla4","Pim1","Ikzf2","Pdcd1","Cd83","Relb","Tnfrsf9","Tnfrsf4"))
# lymphoid tissue
tm_fig2_panel_LT<-rev(c("Stat1","Ccr7","S1pr1","Sell","Bcl2","Tcf7"))
```

### Scatterplot with GOI highlighted (Figure 3 b)
* Hereafter, the scatter plot is generated and plotted.
```{r, fig.height=2.6, fig.width=2.2, out.width="20%", warning=FALSE}
cvnc.pc.scat.colon.tm <- ggplot(cvnc.pc.tr.qstat.scat.filt) + 
  aes(x=logCPMnonColon, y=logCPMcolon) +
  scale_x_continuous(limits = c(-2,14),
                     breaks = c(0,3,6,9,12),
                     expand = c(0,0)
                     )+
  scale_y_continuous(limits = c(-2,14),
                     breaks = c(0,3,6,9,12),
                     expand = c(0,0)
                     )+
  geom_point(colour=rgb(50,50,50,30,maxColorValue = 255), shape=16, size = .375) +
  geom_point(data = cvnc.pc.tr.qstat.scat.filt.UP, aes(x=logCPMnonColon, y=logCPMcolon), colour="sienna", shape=16, size = .5, alpha=.75) +
  geom_point(data = cvnc.pc.tr.qstat.scat.filt.DN, aes(x=logCPMnonColon, y=logCPMcolon), colour="slateblue", shape=16, size = .5, alpha=.75) +
geom_text_repel(data = subset(cvnc.pc.tr.qstat.scat.filt, direction = "both",
                                GeneSymbol %in% tm_fig2_panel_NLT),aes(label=GeneSymbol),
                  force = 50,
                  force_pull = 1,
                  size=1.66,
                  min.segment.length=0,
                  segment.size=0.15,
                  segment.color = "brown4",
                  point.padding=unit(0.00,"cm"),
                  box.padding=unit(0.1,"cm"),
                  max.overlaps = 25,
                  nudge_x = -1,
                  direction = "both",
                  color = "brown4",
                  fontface="italic")+
  geom_text_repel(data = subset(cvnc.pc.tr.qstat.scat.filt, direction = "both",
                                GeneSymbol %in% tm_fig2_panel_LT),aes(label=GeneSymbol),
                  force = 50,
                  force_pull = 1,
                  size=1.66,
                  min.segment.length=0,
                  segment.size=0.15,
                  segment.color = "navy",
                  point.padding=unit(0.00,"cm"),
                  box.padding=unit(0.1,"cm"),
                  max.overlaps = 25,
                  nudge_x = 1,
                  direction = "both",
                  color = "navy",
                  fontface="italic") +
  ggtitle(label = "DEGs colon vs. non-colon", subtitle = "transplanted Treg \nNLT and LT from Miragaia et al. 2019") +
  xlab(bquote(log[2]~CPM *"("*spleen*","*liver*")"))+
  ylab(bquote(log[2]~CPM *"("*colon*")"))+
  theme(plot.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 6, face = "bold"),
        axis.title = element_text(size = 8),
        axis.text.x = element_text(size = 6, colour = "black"),
        axis.text.y =element_text(size = 6, colour = "black"),
        axis.ticks.length.x = unit(0.1,"cm"),
        axis.ticks.length.y = unit(0.1,"cm"),
        axis.ticks = element_line(size=.236),
        plot.background = element_rect(fill="white",colour = "white"),
        panel.background = element_rect(fill = NA,color = "black", size=.236),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
cvnc.pc.scat.colon.tm
```

* The scatterplot is also exported into the figure directory.

```{r, warning=FALSE}
pdf(file="./FIGDIR/Figure3b_TregDGEScatter_cvnc.pc_tmGenes.pdf", height=2.6, width=2.2)
cvnc.pc.scat.colon.tm
dev.off()
```

***

# GSEA with baseline colon signatures
* Please not, you should have run "Figure3_TregDGEScatter_GSEA_Heatmap_Metascape_cvnc.b.Cd62Lneg.Rmd" before proceeding.
* The question addressed here is, whether differential genes in organ-resident Treg also tend to be differentially expressed in transplanted Treg.

## Reading in DGE from Baseline Colon Treg
* Data from analogous baseline Treg analysis is read in, hereafter. 
```{r}
bc.up <- read.delim(file = "./ANALDIR/SupplementaryFigure3g_cluster4.colon.up.GeneIDs.txt")
bc.dn <- read.delim(file = "./ANALDIR/SupplementaryFigure3g_cluster1.colon.down.GeneIDs.txt")                  
```

* First, indices have to be generated for the underlying generalized linear model of this analysis.
* Using the fry() function, a rotation gene set testing for linear models is carried out.

```{r}
ind1 <- rownames(cvnc.pc.fit) %in% bc.up[[1]]
ind2 <- rownames(cvnc.pc.fit) %in% bc.dn[[1]]
ind <- list(ind1,ind2)
names(ind) <- c("up in colon","down in colon")
gse.cvnc.pc.treat.lfc0.6 <- fry(dgel.cvnc.pc, index=ind, design=cvnc.pc.design.dge ,contrast=cvnc.pc.con)
gse.cvnc.pc.treat.lfc0.6
```

## GSEA Barcode plot (Figure 3 c) 
* Herafter, the actual barcode plot is generated and visualized
```{r, fig.height=5, fig.width=5}
barcodeplot(cvnc.pc.tr$table$logFC, 
            index=ind1, 
            index2=ind2,
            labels=c("down in colon","up in colon"), 
            xlab = bquote(log[2]*"FC"),
              col.bars = c("peru","lightslateblue"))
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(-5,5,"Brown bars: Genes up in resident Treg in colon", adj = c(0,.5))
text(-5,4,expression(""*italic(P)*"adj.< 4.7x10"^-40), adj = c(0,.5))
text(5,-3.8,"Blue bars:Genes down resident Treg in colon", adj = c(1,.5))
text(5,-4.8,expression(""*italic(P)*"adj.< 1.9x10"^-37), adj = c(1,.5))
```

* There is a large overlap between resident and BMT samples: Hence transplanted Treg acquire organ phenotype rapidly.
* The barcode plot is exported, hereafter.

```{r}
pdf(file="./FIGDIR/Figure3c_cvnc.pc.Barcodeplot_residentColonSignatures.pdf", height=5, width=5)
barcodeplot(cvnc.pc.tr$table$logFC, 
            index=ind1, 
            index2=ind2,
            labels=c("down in colon","up in colon"), 
            xlab = bquote(log[2]*"FC"),
              col.bars = c("peru","lightslateblue")
            )
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(-5,5,"Brown bars: Genes up in resident Treg in colon", adj = c(0,.5))
text(-5,4,expression(""*italic(P)*"adj.< 4.7x10"^-40), adj = c(0,.5))
text(5,-3.8,"Blue bars: Genes down resident Treg in colon", adj = c(1,.5))
text(5,-4.8,expression(""*italic(P)*"adj.< 1.9x10"^-37), adj = c(1,.5))
dev.off()
```

***

# Additional DGE Analysis on the same Subset of samples
* On the same subset of samples (all Treg re-isolated from tissues following transplant), a DGE analysis is performed.
* The comparison of "allo" vs. "poly" Treg is addressed here.

## DGE List Object
* For the purpose of this analysis we require an additional DGEList object.
* A fitting group variable is defined to do so.
```{r}
group.avp.pc<-factor(as.character(metad.cvnc.pc$grp))
dgel.avp.pc<-DGEList(counts = counts.cvnc.pc, group = group.avp.pc, genes = genes.df)
head(dgel.avp.pc$samples, n=4L)
head(dgel.avp.pc$genes, n=4L)
```

## Exclude Rare Transcripts
* The "filterByExpr()" function is used to determine which genes have sufficiently large counts to be retained in a statistical analysis. 
* Only genes fulfilling that criterion are kept and also normalization factors are calculated.
```{r}
keep.avp.pc <- filterByExpr(dgel.avp.pc, group = group.avp.pc)
dgel.avp.pc <- dgel.avp.pc[keep.avp.pc, , keep.lib.sizes=FALSE]
dgel.avp.pc <- calcNormFactors(dgel.avp.pc)
print(paste(table(keep.avp.pc)[[2]],"genes remain in the analysis after filtering!"))
```

***

# Differential Gene Expression (DGE) analysis
* Differential Gene Expression (DGE) is performed in the following part.

## Design for DGE
In the design for DGE it is critical to include the following factors:

* treat: allo/ poly, which is our contrast of interest
* appl: prophylaxis/control as blocking factor
* org: colon/liver/spleen as blocking factors
* pcr2: this is the only applicable technical variable for batch-correction and should also be included as blocking factor

```{r}
treat<-factor(metad.cvnc.pc$treat)
appl<-factor(metad.cvnc.pc$appl)
org<-factor(metad.cvnc.pc$org)
pcr2<-factor(metad.cvnc.pc$PCR2.Cy)
avp.pc.design.dge<-model.matrix(~0+treat+appl+org+pcr2)
colnames(avp.pc.design.dge)<- c("allo","poly","prophylaxis","liver","spleen","twelve","thirteen","fourteen")
as_tibble(avp.pc.design.dge)
```

## Dispersion Estimation
* Hereafter, dispersion is estimated.
```{r}
dgel.avp.pc<-estimateDisp(dgel.avp.pc,design=avp.pc.design.dge,robust=TRUE)
dgel.avp.pc$common.dispersion
```

## GLM Fitting
* Fitting gene wise glm (generalized linear model) for the constructed design.
```{r}
avp.pc.fit<-glmQLFit(dgel.avp.pc,avp.pc.design.dge)
```

## DGE for Contrast of Interest (COI)
* The difference between allo and poly Treg afer transplantation and re-isolation is of interest, hereafter.
* Thus, our contrast of interest is "allo - poly" for the DGE analysis. 
* We can have a quick look at how many genes are DE with a FC > 1.5 and FDR < 0.05.
```{r}
#Contrast of intrest
avp.pc.con<-makeContrasts(allo - poly, levels=avp.pc.design.dge)
# DGE testing against fc = 1.5
avp.pc.tr <- glmTreat(avp.pc.fit, contrast=avp.pc.con, lfc=log2(1.5))
avp.pc.tr.qstat<-topTags(avp.pc.tr,n = Inf)
avp.pc.tr.qdt<-decideTestsDGE(avp.pc.tr)
summary(avp.pc.tr.qdt)
```

### Exporting DGE Statistics
* The data frame contained in the glmTreat function results is stored for direct access.
```{r}
write.table(x = avp.pc.tr.qstat$table,
            file = "./ANALDIR/SupplementaryFigure3a_TregDGE_avp_pro.prc_FC1.5.txt",
            quote=FALSE, 
            sep = "\t",
            col.names = TRUE,
            row.names = FALSE)
```

***

# DGE Visualization
* Hereafter, DEGs are visualized *via* a scatter plot.

## Data Preparation
* The data is filtered based on logCPM and logRPKM thresholds.
* Necessary data transformations for the scatter plot are carried out as well: columns with individual logCPMs (calculated from logCPM and logFC)
```{r}
avp.pc.tr.qstat.scat<-avp.pc.tr.qstat$table #is a data frame
avp.pc.tr.qstat.scat$logRPKM <- avp.pc.tr.qstat.scat$logCPM - log2(avp.pc.tr.qstat.scat$Length * 0.001)
# Commonly used filter criteria  
avp.pc.tr.qstat.scat.filt<-subset(avp.pc.tr.qstat.scat, logCPM > 1 & logRPKM > 1)
nrow(avp.pc.tr.qstat.scat.filt)
# Transformation for scatter
avp.pc.tr.qstat.scat.filt$logCPMpoly<-avp.pc.tr.qstat.scat.filt$logCPM - (avp.pc.tr.qstat.scat.filt$logFC/2) # poly
avp.pc.tr.qstat.scat.filt$logCPMallo<-avp.pc.tr.qstat.scat.filt$logCPM + (avp.pc.tr.qstat.scat.filt$logFC/2) # allo
# Separation of up and down genes
avp.pc.tr.qstat.scat.filt.UP<-subset(avp.pc.tr.qstat.scat.filt, logFC >= 0 & FDR < 0.05)
avp.pc.tr.qstat.scat.filt.DN<-subset(avp.pc.tr.qstat.scat.filt, logFC <= 0 & FDR < 0.05)
nrow(avp.pc.tr.qstat.scat.filt.UP)
nrow(avp.pc.tr.qstat.scat.filt.DN)
```

## Export UP and DN regulated Genes
* Significantly up and down regulated genes are exported in the following.
```{r}
#extract filtered gene lists
write.table(rownames(avp.pc.tr.qstat.scat.filt.UP),
            file = "./ANALDIR/SupplementaryFigure3a_TregDGE_avp_pro.prc.filt.sig.UP.txt",
            col.names = "GeneID",
            quote=FALSE, 
            row.names = FALSE)
write.table(rownames(avp.pc.tr.qstat.scat.filt.DN),
            file = "./ANALDIR/SupplementaryFigure3a_TregDGE_avp_pro.prc.filt.sig.DN.txt",
            col.names = "GeneID",
            quote=FALSE, 
            row.names = FALSE)
```

## Scatterplot

### Definition of Genes of GOI
* Genes of interest defining Treg in general are defined in this step.
* Approach: Intersection of "Treg-meta-signature" and "Immnav-Treg-signature" by Aubert et al. 2020, extended with genes Itgb7 and Sell.
```{r}
# Reading in the intersection of "Treg-meta-signature" and "Immnav-Treg-signature" published by Aubert et al. 2020
aubert.treg.intersect<-readRDS(file = file.path(getwd(),"PUBDATDIR/aubert.treg.intersect.rds"))
# Extension of the GOI by Itgb7 and Sell
aubert.treg.intersect<-c(aubert.treg.intersect,"Itgb7","Sell")
aubert.treg.intersect
```

### Scatterplot with GOI highlighted (Supplementary Figures 3 a)
* Treg signature genes highlighted
```{r, fig.height=2.6, fig.width=2.2, out.width="20%"}
avp.pc.scat <- ggplot(avp.pc.tr.qstat.scat.filt) + 
  aes(x=logCPMpoly, y=logCPMallo) +
  scale_x_continuous(limits = c(-1,16),
                     breaks = c(0,3,6,9,12,15),
                     expand = c(0,0)
                     )+
  scale_y_continuous(limits = c(-1,16),
                     breaks = c(0,3,6,9,12,15),
                     expand = c(0,0)
                     )+
  geom_point(colour=rgb(50,50,50,30,maxColorValue = 255), shape=16, size = .375) +
  geom_point(data = avp.pc.tr.qstat.scat.filt.DN, aes(x=logCPMpoly, y=logCPMallo), colour="steelblue", shape=16, size = .5, alpha=1) +
  geom_point(data = avp.pc.tr.qstat.scat.filt.UP, aes(x=logCPMpoly, y=logCPMallo), colour="orange", shape=16, size = .5, alpha=1) +
  geom_text_repel(data = subset(avp.pc.tr.qstat.scat.filt, direction = "both",
                                GeneSymbol %in% aubert.treg.intersect),aes(label=GeneSymbol),
                  force = 50,
                  force_pull = 1,
                  size=2, 
                  min.segment.length=0, 
                  segment.size=0.15, 
                  point.padding=unit(0.00,"cm"),
                  box.padding=unit(0.1,"cm"),
                  max.overlaps = 25,
                  color = "black",
                  #nudge_x = -4,
                  direction = "both",
                  fontface="italic")+
  ggtitle(label = "DEGs allo vs. poly Treg", subtitle = "in vitro expanded Treg \nTreg signature Aubert et al. & Itgb7, Sell") +
  ylab(bquote(CPM *"("*log[2]*")" *"("*allo~Treg*")"))+
  xlab(bquote(CPM *"("*log[2]*")" *"("*poly~Treg*")"))+
  theme(plot.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 6, face = "bold"),
        axis.title = element_text(size = 8),
        axis.text.x = element_text(size = 6, colour = "black"), #,angle = 90
        axis.text.y =element_text(size = 6, colour = "black"),
        axis.ticks.length.x = unit(0.1,"cm"),
        axis.ticks.length.y = unit(0.1,"cm"),
        axis.ticks = element_line(size=.236),
        plot.background = element_rect(fill="white",colour = "white"),
        panel.background = element_rect(fill = NA,color = "black", size=.236),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
avp.pc.scat
```

* Herafter, the scatterplot is exported into the figure directory.

```{r}
pdf(file="./FIGDIR/SupplementaryFigure3a_TregDGEScatter_avp.pc_TregSignature.pdf", height=2.6, width=2.2)
avp.pc.scat
dev.off()
```

***

# GSEA with *in vitro* allo up Gene Set

## Reading in DGE from *in vitro* allo/poly comparison and GSEA
* GeneID lists of filtered and significantly up/down regulated genes in *in vitro* allo vs. poly Treg are generated in "Figure1_TregDGEScatter_GSEA_avp_invitro.Rmd".
* The results are read in and indices are generated for the current model of "allo versus poly in all BMT Treg".
* Using the fry() function, a rotation gene set testing for linear models is carried out.
```{r}
avp.ivt.tr.qstat.scat.filt.GeneID.UP <- read.delim(file = "./ANALDIR/Figure1e_TregDGE_avp_invitro.filt.sig.UP.txt")
avp.ivt.tr.qstat.scat.filt.GeneID.DN <- read.delim(file = "./ANALDIR/Figure1e_TregDGE_avp_invitro.filt.sig.DN.txt")
ind1 <- rownames(avp.pc.fit) %in% avp.ivt.tr.qstat.scat.filt.GeneID.UP[[1]]
ind2 <- rownames(avp.pc.fit) %in% avp.ivt.tr.qstat.scat.filt.GeneID.DN[[1]]
ind <- list(ind1,ind2)
names(ind) <- c("up in allo","up poly")
gse.avp.pc <- fry(dgel.avp.pc, index=ind, design=avp.pc.design.dge,contrast=avp.pc.con)
gse.avp.pc
```

## GSEA Barcode plot (Supplementary Figure 3 b)
* Barcode plot presenting logFC (allo vs. poly transplanted & re-isolated Treg) ranked Genes that are upregulated in allo *in vitro* expanded Treg.
```{r, fig.height=3.2, fig.width=6}
barcodeplot(avp.pc.tr$table$logFC, 
            index=ind1, 
            labels=c("poly","allo"), 
            xlab = bquote(log[2]*"FC"),
              col.bars = "orange")

par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(-5,4.5,"Orange bars: Genes up in allo expanded donor Treg", adj = c(0,.5))
text(-5,3,expression(italic(P)[adj.] *"<"* ~3.5*"x10"^-15), adj = c(0,.5))

```

* Barcode plot is also exported.

```{r}
pdf(file="./FIGDIR/SupplementaryFigure3b_avp.cp.Barcodeplot_GenesUpInIvtAllo.pdf", height=3.2, width=6)
barcodeplot(avp.pc.tr$table$logFC, 
            index=ind1, 
            labels=c("poly","allo"), 
            xlab = bquote(log[2]*"FC"),
              col.bars = "orange")
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(-5,4.5,"Orange bars:Genes up in allo expanded donor Treg", adj = c(0,.5))
text(-5,2.8,expression(italic(P)[adj.] *"<"* ~3.5*"x10"^-15), adj = c(0,.5))
dev.off()
```

## Summary of Condition/Organ-wise Comparisons
* First, make sure the "Figure3_GSEA_prc.pro_avp.c.l.s.indiv.Rmd" has been run.
* The file "SupplementaryFigure3b_avp_prc.pro_c.l.s_GSEA_ivt.txt" has to be generated first.
* The individual FDR values for the GSEA are visualized as horizontal bar plot.
```{r, fig.height=1.8, fig.width=2.6}
res_pvalues_ivt<-read.table(file = "./ANALDIR/SupplementaryFigure3b_avp_prc.pro_c.l.s_GSEA_ivt.txt", header = TRUE)
res_pvalues_ivt$logFDR <- log10(res_pvalues_ivt$pivt)
res_pvalues_ivt.bar <- ggplot(res_pvalues_ivt) + 
  aes(x=samples, y=-logFDR ) + coord_flip() + 
  scale_x_discrete(limits=rev) +
  scale_y_continuous(expand = c(0,0.1)) +
  geom_bar(stat="identity", fill="orange") +
  ggtitle(label = "GSEA P values", subtitle = "individual allo-poly comparisons") +
  geom_hline(yintercept=-log10(0.05),linetype='dashed',color='red') +
  ylab(bquote(-log[10]*" "*italic(P)[adj.]))+
  xlab(c("GvHD  no GvHD")) +
  theme(plot.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 6, face = "bold"),
        axis.title = element_text(size = 8),
        axis.text.x = element_text(size = 6, colour = "black"), 
        axis.text.y =element_text(size = 6, colour = "black"),
        axis.ticks.length.x = unit(0.1,"cm"),
        axis.ticks.length.y = unit(0.1,"cm"),
        axis.ticks = element_line(size=.236),
        plot.background = element_rect(fill="white",colour = "white"),
        panel.background = element_rect(fill ="white",color = NA),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
res_pvalues_ivt.bar
```

* The summary bar plot is also exported into the figure directory.

```{r}
pdf(file= "./FIGDIR/SupplementaryFigure3b_avp.org.appl.wise.pvalues.bar_GenesUpInIvtAllo.pdf", height=1.8, width=2.6)
res_pvalues_ivt.bar
dev.off()
```

***

# GSEA with camera

## Pathway enrichment analysis for Hallmark gene sets

* Using camera, we look whether pathways from hallmark collections are enriched in our regulated genes.

```{r}
#hallmark
Mm.H <- readRDS(url("http://bioinf.wehi.edu.au/MSigDB/v7.1/Mm.h.all.v7.1.entrez.rds"))
H.ind.cvnc.pc <- ids2indices(Mm.H, id=dgel.avp.pc$genes$EntrezID)
gst.camera.cvnc.pc <- camera.DGEList(dgel.avp.pc,index=H.ind.cvnc.pc,design=avp.pc.design.dge,contrast=avp.pc.con,inter.gene.cor=0.05)
gst.camera.cvnc.pc
```

## GSEA Barcode Plot for Interferon Alpha Response Hallmark Gene Set (Supplementary Figure 3 d)
* The top Hallmark gene set found enriched is presented as barcode plot, hereafter.  
```{r, fig.height=3.2, fig.width=6}
#hallmark HALLMARK_INTERFERON_ALPHA_RESPONSE 
barcodeplot(avp.pc.tr$table$logFC,
            index=H.ind.cvnc.pc[["HALLMARK_INTERFERON_ALPHA_RESPONSE"]], 
            xlab = bquote(log[2]*"FC"),
            labels=c("poly","allo"), 
            col.bars = "steelblue",
            main="HALLMARK_INTERFERON_ALPHA_RESPONSE")
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,4.5,"Steelblue bars: HALLMARK_INTERFERON_ALPHA_RESPONSE", adj = c(1,.5))
text(5,2.8,expression(italic(P)[adj.] *"<"* ~1.9*"x10"^-4), adj = c(1,.5))
```

* Barcode plot is exported, hereafter.

```{r}
pdf(file= "./FIGDIR/SupplementaryFigure3d_avp.pc.Barcodeplot_HALL_IFN.pdf", height=3.2, width=6)
barcodeplot(avp.pc.tr$table$logFC,
            index=H.ind.cvnc.pc[["HALLMARK_INTERFERON_ALPHA_RESPONSE"]], 
            xlab = bquote(log[2]*"FC"),
            labels=c("poly","allo"), 
            col.bars = "steelblue")
par(new=TRUE)
plot.new( )
plot.window( xlim=c(-5,5), ylim=c(-5,5) )
text(5,4.5,"HALLMARK_INTERFERON_ALPHA_RESPONSE", adj = c(1,.5))
text(5,2.8,expression(italic(P)[adj.] *"<"* ~1.9*"x10"^-4), adj = c(1,.5))
dev.off()
```

## Summary of Condition/Organ-wise Comparisons
* First, make sure the "Figure3_GSEA_prc.pro_avp.c.l.s.indiv.Rmd" has been run.
* The file "SupplementaryFigure3b_avp_prc.pro_c.l.s_GSEA_ifn.txt" has to be generated first.
* The individual FDR values for the GSEA are visualized as horizontal bar plot.
```{r, fig.height=1.9, fig.width=2.6}
res_pvalues_ifn<-read.table(file = "./ANALDIR/SupplementaryFigure3b_avp_prc.pro_c.l.s_GSEA_ifn.txt", header = TRUE)
res_pvalues_ifn$logFDR <- log10(res_pvalues_ifn$pifn)
res_pvalues_ifn.bar <- ggplot(res_pvalues_ifn) + 
  aes(x=samples, y=-logFDR ) + coord_flip() + 
  scale_x_discrete(limits=rev) +
  scale_y_continuous(expand = c(0,0.1)) +
  geom_bar(stat="identity", fill="steelblue") +
  ggtitle(label = "GSEA P values", subtitle = "individual allo-poly comparisons") +
  geom_hline(yintercept=-log10(0.05),linetype='dashed',color='red') +
  ylab(bquote(-log[10]*" "*italic(P)[adj.]))+
  xlab(c("GvHD  no GvHD")) +
  theme(plot.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 6, face = "bold"),
        axis.title = element_text(size = 8),
        axis.text.x = element_text(size = 6, colour = "black"), 
        axis.text.y =element_text(size = 6, colour = "black"),
        axis.ticks.length.x = unit(0.1,"cm"),
        axis.ticks.length.y = unit(0.1,"cm"),
        axis.ticks = element_line(size=.236),
        plot.background = element_rect(fill="white",colour = "white"),
        panel.background = element_rect(fill ="white",color = NA),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
res_pvalues_ifn.bar
```

* The summary adj. p value bar plot is exported, hereafter.

```{r}
pdf(file= "./FIGDIR/SupplementaryFigure3d_avp.org.appl.wise.pvalues.bar_HALL_IFN.pdf", height=1.9, width=2.6)
res_pvalues_ifn.bar
dev.off()
```

***

# SessionInfo
```{r, echo=FALSE, warning=FALSE}
sessionInfo(package = NULL)
writeLines(capture.output(sessionInfo(package = NULL)), "./Rscripts/Figure3_TregDGEScatter_GSEA_cvnc.pc_BMT_SessionInfo.txt")
```

***

# Bibliography