---
title: "Treg Signature Genes in freshly isolated, allo and poly expanded Treg - Heatmap visualizing Expression Levels"
subtitle: "Log2-scaled reads per kilobase per million (RPKM) based on group average values on the unlogged scale are represented in a heatmap."
author: "David Dittmar"
date: "`r Sys.Date()`"
bibliography: Rscripts.bib
nocite: '@Aubert638072'
output:
  html_document:
    df_print: paged
    fig_caption: yes
    number_sections: no
    toc: yes
    toc_float: yes
    code_folding: hide 
---

***

```{r, include=FALSE}
# Load required libraries
library(edgeR)
library(AnnotationDbi)
library(org.Mm.eg.db) #for mouse
library(gplots)
```

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
# ...packages.bib file is also written here
knitr::write_bib(.packages(), "Figure1_TregSigHeatmap_fresh.allo.poly_packages.bib")
```

```{r}
# Load required counts, metadata and annotaion RDS files
counts.m.rm<-readRDS("./RCTDIR/counts.m.rm.2021-06-09.rds")
metad.m.rm<-readRDS("./METADIR/metad.m.rm.2021-06-09.rds")
stid<-readRDS("./METADIR/stid_2020-12-21.rds")
```
* First, required data objects (counts, metadata, short transcript ID) and R packages are loaded. 
* This is the starting point for most analyses conducted on bulk RNA-seq data for the associated manuscript.

***

# Data Preparation
* For the analysis, the data and metadata have to be prepared and properly annotated.

## Annotation Data Frame
* An annotation data frame with 4 columns is constructed.
* The additional EntrezID column is required for possible downstream pathway analyses or annotation operations.
```{r}
genes.df<-as.data.frame(strsplit2(stid,"[$]"))
colnames(genes.df)<-c("EnsemblID","GeneSymbol","Length","GeneType")
genes.df$EnsemblID<-as.character(genes.df$EnsemblID)
genes.df$GeneSymbol<-as.character(genes.df$GeneSymbol)
genes.df$Length<-as.numeric(as.character(genes.df$Length))
genes.df$GeneType<-as.character(genes.df$GeneType)
genes.df$EntrezID<- mapIds(org.Mm.eg.db,
                    keys=genes.df$GeneSymbol,
                    column="ENTREZID",
                    keytype="SYMBOL",
                    multiVals="first")
rownames(genes.df)<-stid
head(genes.df, n=5L)
#See if the gene length values are still as they should, so this should yield an empty vector
which(strsplit2(stid,"[$]")[,3] %in% genes.df$Length==FALSE)
```

## Subsetting Index
* In this analysis, we focus on *in vitro* expanded Treg (allo and poly) as well as freshly isolated Treg.
* A subsetting index is generated accordingly.
```{r}
iv.ind<-which(grepl("wt_S_pos",metad.m.rm$grp)|(grepl("d",metad.m.rm$org)&(grepl(c(1),metad.m.rm$tx198_No)|grepl(c(2),metad.m.rm$tx198_No)|grepl(c(4),metad.m.rm$tx198_No)|grepl(c(5),metad.m.rm$tx198_No)|grepl(c(10),metad.m.rm$tx198_No)|grepl(c(12),metad.m.rm$tx198_No))))
iv.ind
```

## Counts
* Here, the comprehensive counts data frame is reduced using the subsetting index.
```{r}
#Subsetting the respective counts object (counts.m.rm)
counts.iv<-counts.m.rm[,iv.ind]
ncol(counts.iv)
```

## Metadata
* The metadata has to be reduced accordingly. 
* The original order of samples is introduced for downstream re-ordering of the read count table.
* A graspable group variable is introduced, as well.
```{r, echo= FALSE}
metad.iv<-metad.m.rm[iv.ind,]
metad.iv$orig.ord<-1:nrow(metad.iv)
metad.iv<-metad.iv[order(metad.iv$org),]
metad.iv$group<-c(rep(c("allo","poly"),6),rep("fresh",4))
counts.ind<-metad.iv$orig.ord
metad.iv
#the group variable for all later assignments of colors/pchs.
group.iv<-factor(as.character(metad.iv$group), levels = c("fresh", "allo",  "poly"))
group.iv
```

## Re-ordering Counts Data Frame Columns
* Here, read counts are re-ordered according to the metadata object.
```{r}
counts.iv.ord<-counts.iv[,counts.ind]
head(counts.iv.ord)
counts.iv<-counts.iv.ord
rm(counts.iv.ord)
```

## Consistency Check
* In order to avoid wrongful labelling, sample_IDs are used to confirm that the counts and metadata objects are consistent.
```{r}
if (sum(colnames(counts.iv)==metad.iv$sample_ID)==length(iv.ind)) {
  print("Counts and metadata are properly subsetted and in accordance with each other. The analysis can be proceeded with")
}else {
  print("Something is inconsistent, check the code above for errors during subsetting!")
}
```

## DGE List Object
* A DGEList object is generated, the main class of the edgeR package. 
* For full functionality, the counts.ivt, group.avp.ivt and genes.df variables are used to do so.
```{r}
dgel.iv<-DGEList(counts = counts.iv, group = group.iv, genes = genes.df)
head(dgel.iv$counts, n=4L)
head(dgel.iv$samples, n=4L)
head(dgel.iv$genes, n=4L)
```

## Exclusion of Rare Transcripts
* The "filterByExpr()" function is used to determine which genes have sufficiently large counts to be retained in a statistical analysis. 
* Only genes fulfilling that criterion are kept and also normalization factors are calculated.
```{r}
keep.iv <- filterByExpr(dgel.iv)
dgel.iv <- dgel.iv[keep.iv, , keep.lib.sizes=FALSE]
dgel.iv <- calcNormFactors(dgel.iv)
table(keep.iv)
dgel.iv$samples
```

## Generation of RPKM variables per Group
* Hereafter, RPKM values (log-scaled with prior.count = 2) are computed per group based on group average values on the unlogged scale.
* Using cbind.data.frame(), the GeneSymbol column is added to the dataframe, preventing conversion to character columns.
```{r}
# log rpkmByGroup
iv.log.rpkmByGroup<-rpkmByGroup(dgel.iv, normalized.lib.sizes = TRUE, log = TRUE, prior.count = 2)
# GeneSymbol column
iv.log.rpkmByGroup<-cbind.data.frame(iv.log.rpkmByGroup, GeneSymbol=strsplit2(rownames(iv.log.rpkmByGroup), split= "$", fixed = T)[,2])
```

***

# Log-scaled RPKM Heatmap representing Genes of Interest (GOI)

## Definition of Genes of GOI
* Genes of interest defining Treg in general are defined in this step.
* Approach: Intersection of "Treg-meta-signature" and "Immnav-Treg-signature" by Aubert et al. 2020, extended with genes Itgb7 and Sell.
```{r}
# Reading in the intersection of "Treg-meta-signature" and "Immnav-Treg-signature" published by Aubert et al. 2020
aubert.treg.intersect<-readRDS(file = file.path(getwd(),"PUBDATDIR/aubert.treg.intersect.rds"))
# Extension of the GOI by Itgb7 and Sell
aubert.treg.intersect<-c(aubert.treg.intersect,"Itgb7","Sell")
aubert.treg.intersect
```
* Also export GOI as tab-delimited text file.
```{r}
write.table(x = aubert.treg.intersect,
            file = "./PUBDATDIR/aubert.treg.intersect.extended.txt",
            col.names = "GeneSymbol",
            quote = FALSE,
            row.names = FALSE)
```

## Heatmap with log-scaled RPKM for GOI (Figure 1 f)
* A pre-filtering for log2(RPKM) values greater than 1 for either allo or poly Treg is performed.
* Genes remaining in the analysis and intersecting with the extended Treg Signature by Aubert et al. 2020 are represented as a heatmap.
```{r, fig.height=5, fig.width=3.2, warning=FALSE}
# Pre-filtering and GOI intersection
iv.log.rpkmByGroup.filtered<-subset(iv.log.rpkmByGroup, allo > 1 | poly > 1)
ind.aubert<-iv.log.rpkmByGroup.filtered$GeneSymbol%in%aubert.treg.intersect
aubert.hm.matrix<-data.matrix(iv.log.rpkmByGroup.filtered[ind.aubert,1:3])
# The actual heatmap.2 plot command
# Orientation for plot elements: (1=heatmap, 2=row dendogram, 3=col dendogram, 4= key )
heatmap.2(x = aubert.hm.matrix,
          Rowv=T,
          Colv=NA,
          distfun = function(c){dist(c, method = "manhattan")},
          hclustfun=function(c){hclust(c, method="ward.D")},
          dendrogram = "row",
          symm = FALSE,
          na.rm=TRUE,
          col = hcl.colors(30, palette = "ag_Sunset", rev=F),
          margins=c(1.8,2.5),
          cexRow = 1/log10(nrow(aubert.hm.matrix)),
          labRow = as.expression(lapply(iv.log.rpkmByGroup.filtered[ind.aubert,4],function(a) bquote(italic(.(a))))),
          srtRow = 0,
          offsetRow = 0,
          labCol=as.expression(lapply(paste0(colnames(aubert.hm.matrix),c("\n(n=4)","\n(n=6)","\n(n=6)")), function(a) bquote(italic(.(a))))),
          srtCol = 0,
          offsetCol = 0,
          adjCol = c(0.45,1.5),
          #cexCol = 1,
          cexCol=1,
          key=T,
          symbreaks = FALSE,
          symkey = FALSE,
          keysize = 0.005,
          key.par = list(mai=c(0.1,0.1,0.1,0.1)),
          key.title = "none", 
          key.xlab = expression(log[2]*"("*RPKM*")"),
          density.info="none",
          trace = "none", 
          lmat = matrix(c(4,2,3,1),nrow = 2, ncol = 2),
          lhei=c(1,3),
          lwid = c(0.25, 0.3))
```

* The heatmap is also exported into the figure directory.

```{r message=FALSE, warning=FALSE}
pdf(file = "./FIGDIR/Figure1f_TregSigHeatmap_fresh.allo.poly.pdf",height = 5,width = 3.2)
heatmap.2(x = aubert.hm.matrix,
          Rowv=T,
          Colv=NA,
          distfun = function(c){dist(c, method = "manhattan")},
          hclustfun=function(c){hclust(c, method="ward.D")},
          dendrogram = "row",
          symm = FALSE,
          na.rm=TRUE,
          col = hcl.colors(30, palette = "ag_Sunset", rev=F),
          margins=c(1.8,2.5),
          cexRow = 1/log10(nrow(aubert.hm.matrix)),
          labRow = as.expression(lapply(iv.log.rpkmByGroup.filtered[ind.aubert,4],function(a) bquote(italic(.(a))))),
          srtRow = 0,
          offsetRow = 0,
          labCol=as.expression(lapply(paste0(colnames(aubert.hm.matrix),c("\n(n=4)","\n(n=6)","\n(n=6)")), function(a) bquote(italic(.(a))))),
          srtCol = 0,
          offsetCol = 0,
          adjCol = c(0.45,1.5),
          #cexCol = 1,
          cexCol=1,
          key=T,
          symbreaks = FALSE,
          symkey = FALSE,
          keysize = 0.005,
          key.par = list(mai=c(0.1,0.1,0.1,0.1)),
          key.title = "none", 
          key.xlab = expression(log[2]*"("*RPKM*")"),
          density.info="none",
          trace = "none", 
          lmat = matrix(c(4,2,3,1),nrow = 2, ncol = 2),
          lhei=c(1,3),
          lwid = c(0.25, 0.3))
dev.off()
```

***

# SessionInfo
```{r, echo=FALSE, warning=FALSE}
sessionInfo(package = NULL)
writeLines(capture.output(sessionInfo(package = NULL)), "./Rscripts/Figure1_TregSigHeatmap_fresh.allo.poly_SessionInfo.txt")
```

***

# Bibliography